Reference Tracking Strategies of Deaf Adult Signers in Turkish Sign Language

### 2023-05-04 revised_manuscript.Rmd
updated

In TİD and many other sign languages, it is possible to use bare nouns as in \ref{ex:2a} as well as fingerspelled nouns as in \ref{ex:2b}. Fingerspelling refers to the use of the manual alphabet to spell all or part of the letters of a word borrowed from a surrounding written language [@kubus2008]. Nouns can also be modified by a prenominal or postnominal extended index finger pointing sign, glossed as IX, functioning as a determiner or a demonstrative as in \ref{ex:2c} [@maclaughlin1997] or by a modifying classifier within the same NP as in \ref{ex:2d} [@frederiksen2016].\footnote{Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions (Lillo-Martin and Klima, 1990). In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.}

```{=tex}
\newpage
\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:2a} Full noun \\\glll
{} \includegraphics[width=35pt]{pictures/p1.png} {} \\ \textbf{Gloss:} \textsc{mouse\footnotemark} \textsc{sick} \\ 
\textbf{Status:} Intro. \\
\glt \textbf{Context:} Narration 6, Event 1. Here, the signer introduces the mouse, Jerry, at the beginning of the narration with a bare noun.
\glt \textbf{Translation:} There is a sick mouse.
\vspace{0.35cm}

@@

Following @berman2013relating, we first identified the clause boundaries in each sentence indicated by the presence of predicates and informative non-manual markers (e.g., head nod). On the second tier, we coded the discourse status for each referent. Having adapted the classification of @gullberg2006, we considered all referents mentioned for the first time as introduced regardless of their syntactic function in the sentence. Only subjects were coded for maintenance and re-introduction. We coded subject referents as maintained only if they had been referred to in the previous clause. To this end, the referents which had been discontinued across one or more sentences were all regarded re-introduced into the discourse. On a third tier, we had the RE type for a single reference with the main categories nominal, pronominal, zero anaphora, and classifier. Under nominal, we identified bare nouns, modified nouns, and fingerspelled nouns. For classifiers, we distinguished among three types of handshapes: WECL, BPCL, and ExtCL. Finally, we had CA[^1], agreement, plain tags for zero anaphora. We were unable to observe any instance of HCLs occurring in isolation (i.e., not part of CA). We should also highlight here that we coded referent status as nominal if the clause included an explicit nominal referent. For example, if constructed action, an index sign, or a classifier preceded or followed its explicitly signed nominal referent, the clause was regarded as having nominal reference (i.e., modified noun). The annotation tiers are presented in Table \ref{tab:tab-2}.

[^1]: Note that CA can also be further subdivided into overt, reduced, and subtle CA, following [@cormier_rethinking_2015]. In overt CA, the signer is "fully in character" and does not use any other lexical or classifier constructions. In reduced CA, the signer can involve a small amount of narration or lexical material while tracking a referent using CA. However, subtle CA refers to cases where the referent is primarily tracked by a non-CA construction (e.g., nominal, pronominal, verbal or classifier constructions) but the signer subtly uses a few CA articulators at the same time (e.g., enacting the eye-gaze of the character). We only coded overt and reduced types as reference tracking by CA, and disregarded the subtle type. We'd also like to acknowledge that mixed CA, or simultaneous tracking of two or multiple references through CA and other tools can be possible [@dudis2004; @liddell2003]. In such cases, as we explain here, we only coded only the subject of a clause for maintenance and re-introduction. If the referent was in subject position and tracked by using CA, we considered this an instance of CA, as well.

@@

```{=tex}
\begin{table}
\small
\caption{Distribution of all discourse status by RE types}
\label{tab:my-table}
\begin{tabular}{llll}
\hline
                     & Introduced    & Maintained     & Re-introduced \\ \hline
\textit{Nominal:}    & 92.7\% (956) & 5.9\% (93)    & 63.7\% (884) \\
\textit{Pronominal:} & 0.1\% (1)    & 0.3\% (4)     & 2.5\% (35)   \\
\textit{Classifier:} & 5.3\% (55)   & 18.3\% (291)  & 7.8\% (108)     \\
\textit{Null:}       & 1.8\% (19)   & 75.6\% (1201) & 26\% (360) \\ \hline
\end{tabular}%
\end{table}
```
Table \ref{tab:my-table2} shows that for all discourse statuses the most numerous nominal type was bare nominal which was followed by nouns modified by index or classifiers. In number, signers used modified nouns more in introduction and re-introduction compared to maintenance settings. Signers produced a total number of `r sum(df_narration$ref_type == "Pronominal")` pronouns (Table \ref{tab:my-table3}). A great majority of the pronominal constructions were used to re-introduce a referent into the discourse. We observed only a very small amount of pronoun use for referent introduction and maintenance.

```{=tex}
\begin{table}
\small
\caption{Distribution of nominals by discourse status}
\label{tab:my-table2}
\begin{tabular}{lllll}
\hline
                        & Bare Noun  & Modified Noun & FS Noun \\ \hline
\textit{Introduced:}    & 48.7\% (813) & 53\% (134)     & 75\% (9) \\
\textit{Maintained:}    & 5\% (83)  & 4\% (10)      & 0\% (0) \\
\textit{Re-introduced:} & 46.3\% (772) & 43\% (109)     & 25\% (3) \\ \hline
\end{tabular}
\end{table}

\begin{table}
\small
\caption{Distribution of pronominals by discourse status}
\label{tab:my-table3}
\begin{tabular}{ll}
\hline
                        & Pronominal  \\ \hline
\textit{Introduced:}    & 2.5\% (1)   \\
\textit{Maintained:}    & 10\% (4)    \\
\textit{Re-introduced:} & 87.5\% (35) \\ \hline
\end{tabular}
\end{table}
```
As for the classifier constructions (Table \ref{tab:my-table4}), WECL was the most common type, followed by BPCL and ExtCL. For the first mention of a referent, signers mostly preferred to use ExtCL. However, for referent maintenance and re-introduction WECL was used more compared to BPCL and ExtCL.

Among the markers of zero anaphora, CA was most frequently employed, and it was followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}). Overall, signers used very few markers of zero anaphora for the introduction of a referent and the majority of the existing ones were CA. Additionally, signers used CA more for referent re-introduction than maintenance. We observed similar instances of agreement verbs for maintained and re-introduced contexts. In contrast, plain verbs were favored more in the maintenance setting than re-introduction.

```{=tex}
\begin{table}
\small
\caption{Distribution of classifiers by discourse status}
\label{tab:my-table4}
\begin{tabular}{llll}
\hline
                        & WECL          & BPCL        & ExtCL         \\ \hline
\textit{Introduced:}    & 2.2\% (6)    & 2.5\% (3)   & 70.8\% (46) \\
\textit{Maintained:}    & 67.5\% (181) & 77.7\% (94) & 24.6\% (16) \\
\textit{Re-introduced:} & 30.2\% (81)   & 19.8\% (24)   & 4.6\% (3)   \\ \hline
\end{tabular}
\end{table}
```
Across the discourse statuses, there is also a difference between participant groups in the use of zero anaphora for reference tracking. DoD signers of TİD used it more frequently (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "DoD") / nrow(df_narration), digits=2)`) compared to the DoH signers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "DoH") / nrow(df_narration), digits=2)`). Comparably, DoH signers used nominals more frequently (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "DoH") / nrow(df_narration), digits=2)`) than DoD signers (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "DoD") / nrow(df_narration), digits=2)`). Figure 1 shows the numerical distribution of the responses by RE type and Group.

```{=tex}
\begin{table}
\small
\caption{Distribution of zero anaphora by discourse status}
\label{tab:my-table5}
\begin{tabular}{llll}
\hline
                        & Constructed Action & Plain Verb    & Agreement Verb \\ \hline
\textit{Introduced:}    & 1.6\% (17)        & 0.3\% (1)    & 0.5\% (1)      \\
\textit{Maintained:}    & 71.5\% (749)      & 89.3\% (302) & 76.9\% (150)  \\
\textit{Re-introduced:} & 26.8\% (281)      & 10.4\% (35)   & 22.6\% (44)   \\ \hline
\end{tabular}
\end{table}
```
```{r fig-count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Zero Anaphora), and Participant Group (DoD, DoH)"}
df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "ZERO"))

@@

For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using discourse status and acquisition group as predictors. We used the brms package [@burkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH). As the overall number of produced pronouns was very little (40 data points), we excluded pronominal use from the following analyses. For the current analysis, we did not define priors ourselves and used the preset uninformed priors. Figure 2[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals. There are 3 responses we are considering in this model. They are Classifier, Nominal, and Zero Anaphora. Classifier response is the level that other responses Nominal and Zero Anaphora are compared against.

[^2]: The regression model results are given in \ref{section:modelResult}. We ran another model where we used the self-reported acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1, and 7-above:Late2. However, our model did not converge properly and the credible intervals were very wide spread, yielding the results uninformative. We provide the model results as a plot in \ref{section:d}.

@@

As for the distribution of the zero anaphora, approximately 90% of the plain verbs were used to maintain a previous referent and this makes these constructions least implicit REs used for TİD[^3]. Agreement verbs and CA were the following implicit markers of reference, `r sum(df_narration$GTR1 == "Person agreement verb" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1 == "Person agreement verb")` and `r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1_plot == "CA")` of which were used for maintained contexts, respectively. This contrasts with some of the research [@czubek2017] in which agreement verbs ranked lowest on the scale, indicating that these constructions were more implicit compared to plain verbs. However, there seems to be a common ground for the position of CA markers which have been claimed to rank higher than other markers of zero anaphora as in the present study [@frederiksen2016; @czubek2017; @barbera2009]. In Figure \ref{fig-scale}, we report the distribution of REs in TİD as a function of referent accessibility and compare our accessibility hierarchy with that of [@frederiksen2016] for ASL. In the hierarchies, the leftmost REs (e.g., modified and bare nouns for TİD) are used when the accessibility is low. If the referent is already active in the addressee's mind and its accessibility is high, then the rightmost markers (e.g., agreement and plain verbs for TİD) are selected.

[^3]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping. For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion [@meir1999]. As for plain verbs, both phonology of a sign and its semantics have an effect on morphology. For instance, \textsc{eat} in TİD is signed nearby the mouth and since it needs to remain in that phonological location, it is not suitable for agreement and thus it is regarded as a plain verb. Likewise the experiencer verb \textsc{love} is signed on the torso and as such it does not take part in agreement. These generalizations noted we should also be cautious in labelling a verb as categorially one type or the other. For instance, some intransitive plain verbs like \textsc{fall} or \textsc{jump} in TİD may display spatial person agreement with the single argument in the clause [@gokgoz2020]. See @lourenço2018 also for a similar discussion in Brazilian Sign Language.

>>>>

In TİD and many other sign languages, it is possible to use bare nouns as in \ref{ex:2a} as well as fingerspelled nouns as in \ref{ex:2b}. Fingerspelling refers to the use of the manual alphabet to spell all or part of the letters of a word borrowed from a surrounding written language [@kubus2008]. Nouns can also be modified by a prenominal or postnominal extended index finger pointing sign, glossed as IX, functioning as a determiner or a demonstrative as in \ref{ex:2c} [@maclaughlin1997] or by a modifying classifier within the same NP as in \ref{ex:2d} [@frederiksen2016].[^1]

[^1]: Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions [@lillo-martin1990]. In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.

```{=tex}
\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:2a} Full noun \\\glll
{} \includegraphics[width=35pt]{pictures/p1.png} {} \\ \textbf{Gloss:} \textsc{mouse\footnotemark} \textsc{sick} \\ 
\textbf{Status:} Intro. \\
\glt \textbf{Context:} Narration 6, Event 1. Here, the signer introduces the mouse, Jerry, at the beginning of the narration with a bare noun.
\glt \textbf{Translation:} There is a sick mouse.
\vspace{0.35cm}

@@

Following @berman2013relating, we first identified the clause boundaries in each sentence indicated by the presence of predicates and informative non-manual markers (e.g., head nod). On the second tier, we coded the discourse status for each referent. Having adapted the classification of @gullberg2006, we considered all referents mentioned for the first time as introduced regardless of their syntactic function in the sentence. Only subjects were coded for maintenance and re-introduction. We coded subject referents as maintained only if they had been referred to in the previous clause. To this end, the referents which had been discontinued across one or more sentences were all regarded re-introduced into the discourse. On a third tier, we had the RE type for a single reference with the main categories nominal, pronominal, zero anaphora, and classifier. Under nominal, we identified bare nouns, modified nouns, and fingerspelled nouns. For classifiers, we distinguished among three types of handshapes: WECL, BPCL, and ExtCL. Finally, we had CA[^2], agreement, plain tags for zero anaphora. We were unable to observe any instance of HCLs occurring in isolation (i.e., not part of CA). We should also highlight here that we coded referent status as nominal if the clause included an explicit nominal referent. For example, if constructed action, an index sign, or a classifier preceded or followed its explicitly signed nominal referent, the clause was regarded as having nominal reference (i.e., modified noun). The annotation tiers are presented in Table \ref{tab:tab-2}.

[^2]: Note that CA can also be further subdivided into overt, reduced, and subtle CA, following [@cormier_rethinking_2015]. In overt CA, the signer is "fully in character" and does not use any other lexical or classifier constructions. In reduced CA, the signer can involve a small amount of narration or lexical material while tracking a referent using CA. However, subtle CA refers to cases where the referent is primarily tracked by a non-CA construction (e.g., nominal, pronominal, verbal or classifier constructions) but the signer subtly uses a few CA articulators at the same time (e.g., enacting the eye-gaze of the character). We only coded overt and reduced types as reference tracking by CA, and disregarded the subtle type. We'd also like to acknowledge that mixed CA, or simultaneous tracking of two or multiple references through CA and other tools can be possible [@dudis2004; @liddell2003]. In such cases, as we explain here, we only coded only the subject of a clause for maintenance and re-introduction. If the referent was in subject position and tracked by using CA, we considered this an instance of CA, as well.

@@

Table \ref{tab:my-table2} shows that for all discourse statuses the most numerous nominal type was bare nominal which was followed by nouns modified by index or classifiers. In number, signers used modified nouns more in introduction and re-introduction compared to maintenance settings. Signers produced a total number of `r sum(df_narration$ref_type == "Pronominal")` pronouns (Table \ref{tab:my-table3}). A great majority of the pronominal constructions were used to re-introduce a referent into the discourse. We observed only a very small amount of pronoun use for referent introduction and maintenance.

```{=tex}
\begin{table}
\small
\caption{Distribution of all discourse status by RE types}
\label{tab:my-table}
\begin{tabular}{llll}
\hline
                     & Introduced    & Maintained     & Re-introduced \\ \hline
\textit{Nominal:}    & 92.7\% (956) & 5.9\% (93)    & 63.7\% (884) \\
\textit{Pronominal:} & 0.1\% (1)    & 0.3\% (4)     & 2.5\% (35)   \\
\textit{Classifier:} & 5.3\% (55)   & 18.3\% (291)  & 7.8\% (108)     \\
\textit{Null:}       & 1.8\% (19)   & 75.6\% (1201) & 26\% (360) \\ \hline
\end{tabular}%
\end{table}

\begin{table}
\small
\caption{Distribution of nominals by discourse status}
\label{tab:my-table2}
\begin{tabular}{lllll}
\hline
                        & Bare Noun  & Modified Noun & FS Noun \\ \hline
\textit{Introduced:}    & 48.7\% (813) & 53\% (134)     & 75\% (9) \\
\textit{Maintained:}    & 5\% (83)  & 4\% (10)      & 0\% (0) \\
\textit{Re-introduced:} & 46.3\% (772) & 43\% (109)     & 25\% (3) \\ \hline
\end{tabular}
\end{table}

\begin{table}
\small
\caption{Distribution of pronominals by discourse status}
\label{tab:my-table3}
\begin{tabular}{ll}
\hline
                        & Pronominal  \\ \hline
\textit{Introduced:}    & 2.5\% (1)   \\
\textit{Maintained:}    & 10\% (4)    \\
\textit{Re-introduced:} & 87.5\% (35) \\ \hline
\end{tabular}
\end{table}
```
```{=tex}
\begin{table}
\small
\caption{Distribution of classifiers by discourse status}
\label{tab:my-table4}
\begin{tabular}{llll}
\hline
                        & WECL          & BPCL        & ExtCL         \\ \hline
\textit{Introduced:}    & 2.2\% (6)    & 2.5\% (3)   & 70.8\% (46) \\
\textit{Maintained:}    & 67.5\% (181) & 77.7\% (94) & 24.6\% (16) \\
\textit{Re-introduced:} & 30.2\% (81)   & 19.8\% (24)   & 4.6\% (3)   \\ \hline
\end{tabular}
\end{table}
```
```{=tex}
\begin{table}
\small
\caption{Distribution of zero anaphora by discourse status}
\label{tab:my-table5}
\begin{tabular}{llll}
\hline
                        & Constructed Action & Plain Verb    & Agreement Verb \\ \hline
\textit{Introduced:}    & 1.6\% (17)        & 0.3\% (1)    & 0.5\% (1)      \\
\textit{Maintained:}    & 71.5\% (749)      & 89.3\% (302) & 76.9\% (150)  \\
\textit{Re-introduced:} & 26.8\% (281)      & 10.4\% (35)   & 22.6\% (44)   \\ \hline
\end{tabular}
\end{table}
```
```{r fig-count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Zero Anaphora), and Participant Group (DoD, DoH)"}
df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "ZERO"))

@@

Across the discourse statuses, there is also a difference between participant groups in the use of zero anaphora for reference tracking. DoD signers of TİD used it more frequently (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "DoD") / nrow(df_narration), digits=2)`) compared to the DoH signers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "DoH") / nrow(df_narration), digits=2)`). Comparably, DoH signers used nominals more frequently (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "DoH") / nrow(df_narration), digits=2)`) than DoD signers (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "DoD") / nrow(df_narration), digits=2)`). Figure 1 shows the numerical distribution of the responses by RE type and Group.

For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using discourse status and acquisition group as predictors. We used the brms package [@burkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH). As the overall number of produced pronouns was very little (40 data points), we excluded pronominal use from the following analyses. For the current analysis, we did not define priors ourselves and used the preset uninformed priors. Figure 2[^3] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals. There are 3 responses we are considering in this model. They are Classifier, Nominal, and Zero Anaphora. Classifier response is the level that other responses Nominal and Zero Anaphora are compared against.

[^3]: The regression model results are given in \ref{section:modelResult}. We ran another model where we used the self-reported acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1, and 7-above:Late2. However, our model did not converge properly and the credible intervals were very wide spread, yielding the results uninformative. We provide the model results as a plot in \ref{section:d}.

@@

As for the distribution of the zero anaphora, approximately 90% of the plain verbs were used to maintain a previous referent and this makes these constructions least implicit REs used for TİD[^4]. Agreement verbs and CA were the following implicit markers of reference, `r sum(df_narration$GTR1 == "Person agreement verb" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1 == "Person agreement verb")` and `r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1_plot == "CA")` of which were used for maintained contexts, respectively. This contrasts with some of the research [@czubek2017] in which agreement verbs ranked lowest on the scale, indicating that these constructions were more implicit compared to plain verbs. However, there seems to be a common ground for the position of CA markers which have been claimed to rank higher than other markers of zero anaphora as in the present study [@frederiksen2016; @czubek2017; @barbera2009]. In Figure \ref{fig-scale}, we report the distribution of REs in TİD as a function of referent accessibility and compare our accessibility hierarchy with that of [@frederiksen2016] for ASL. In the hierarchies, the leftmost REs (e.g., modified and bare nouns for TİD) are used when the accessibility is low. If the referent is already active in the addressee's mind and its accessibility is high, then the rightmost markers (e.g., agreement and plain verbs for TİD) are selected.

[^4]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping. For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion [@meir1999]. As for plain verbs, both phonology of a sign and its semantics have an effect on morphology. For instance, \textsc{eat} in TİD is signed nearby the mouth and since it needs to remain in that phonological location, it is not suitable for agreement and thus it is regarded as a plain verb. Likewise the experiencer verb \textsc{love} is signed on the torso and as such it does not take part in agreement. These generalizations noted we should also be cautious in labelling a verb as categorially one type or the other. For instance, some intransitive plain verbs like \textsc{fall} or \textsc{jump} in TİD may display spatial person agreement with the single argument in the clause [@gokgoz2020]. See @lourenço2018 also for a similar discussion in Brazilian Sign Language.

### 2023-05-04 revised_manuscript.tex
updated

\newpage
\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:2a} Full noun \\\glll
{} \includegraphics[width=35pt]{pictures/p1.png} {} \\ \textbf{Gloss:} \textsc{mouse\footnotemark} \textsc{sick} \\ 
\textbf{Status:} Intro. \\
\glt \textbf{Context:} Narration 6, Event 1. Here, the signer introduces the mouse, Jerry, at the beginning of the narration with a bare noun.
\glt \textbf{Translation:} There is a sick mouse.
\vspace{0.35cm}

@@

As for the classifier constructions (Table \ref{tab:my-table4}), WECL
was the most common type, followed by BPCL and ExtCL. For the first
mention of a referent, signers mostly preferred to use ExtCL. However,
for referent maintenance and re-introduction WECL was used more compared
to BPCL and ExtCL.

Among the markers of zero anaphora, CA was most frequently employed, and
it was followed by plain verbs and agreement verbs (Table
\ref{tab:my-table5}). Overall, signers used very few markers of zero
anaphora for the introduction of a referent and the majority of the
existing ones were CA. Additionally, signers used CA more for referent
re-introduction than maintenance. We observed similar instances of
agreement verbs for maintained and re-introduced contexts. In contrast,
plain verbs were favored more in the maintenance setting than
re-introduction.

@@

>>>>

\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:2a} Full noun \\\glll
{} \includegraphics[width=35pt]{pictures/p1.png} {} \\ \textbf{Gloss:} \textsc{mouse\footnotemark} \textsc{sick} \\ 
\textbf{Status:} Intro. \\
\glt \textbf{Context:} Narration 6, Event 1. Here, the signer introduces the mouse, Jerry, at the beginning of the narration with a bare noun.
\glt \textbf{Translation:} There is a sick mouse.
\vspace{0.35cm}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-lillo-martin1990}{}}%
Lillo-Martin, Diane, Klima, Edward S., 1990. Pointing out differences:
ASL pronouns in syntactic theory. Theoretical issues in sign language
research, Volume 1: Linguistics 191.
